---
title: "CDUP"
output: html_document
---

```{r}
setwd("~/Desktop/CDUP")
#setwd("~/Columbia/Commerce_Data_Usability_Project/Maps & Recodes") # Note: change to your working directory
cfs <- read.csv("2012_cfs.txt")
setwd("~/Desktop/Commerce_Data_Usability_Project/")
FIPS.Code <- read.csv("Recode Sources/FIPS Code.csv")
NAICS <- read.csv("Recode Sources/NAICS.csv")
SCTG <- read.csv("Recode Sources/SCTG.csv")
MT <- read.csv("Recode Sources/MT.csv")
```

Shapefile of CFS areas + other info: http://www.census.gov/econ/census/help/geography/cfs_areas.html

Info: http://www.census.gov/econ/cfs/pums.html

# Objective

$$GIS and Network Analysis: mapping centrality and power for different industries, across CFS Areas$$

# Recodes

FIPS Codes:
```{r}
cfs$ORIG_STATE_name <- FIPS.Code$State[match(cfs$ORIG_STATE, FIPS.Code$FIPS)]
cfs$DEST_STATE_name <- FIPS.Code$State[match(cfs$DEST_STATE, FIPS.Code$FIPS)]
```

NAICS (North American Industry Classification System) Codes
```{r}
cfs$NAICS_name <- NAICS$Description[match(cfs$NAICS, NAICS$NAICS)]
```

SCTG (Standard Classification of Transported Goods) Codes
```{r}
cfs$SCTG_name <- SCTG$Description[match(cfs$SCTG, SCTG$SCTG)]
```

MT (Mode of Transportation)
```{r}
cfs$MODE_name <- MT$Mode.of.transportation.Description[match(cfs$MODE, MT$Mode.Code)]
```

## Notes on suppressed data
Metropolitan areas:

```{r}
length(unique(cfs$ORIG_CFS_AREA))
length(unique(cfs$DEST_CFS_AREA))
length(unique(c(cfs$DEST_CFS_AREA, cfs$ORIG_CFS_AREA)))
```

Note: The difference in number of locations is due to ORIG_MA = 00000 code, for MA origin suppressed & CFS AREA = 00-00000 for shipment origin suppressed completely.

```{r}
sum(cfs$ORIG_MA == 0)
```

SCTG suppressed code: 00
Mode suppressed code: 00

Percentage of dollar value suppressed:
Percentage of weight suppressed:
Percentage of hazardous material suppressed:

# Geographic visualizations

In QGIS: CFS.Code (string 8) = "ANSI_ST" + '-' + "CFS12_AREA"
Dissolved by CFS_Code
Note: 132 areas
Join with cvs file below.

Example 1: All products
Group by origin CFS Area. Sum by dollar value & weight
```{r}
library(dplyr)
cfs_by_dest <- group_by(cfs, DEST_CFS_AREA)

Dest_summary <- summarize(cfs_by_dest, value = sum(SHIPMT_VALUE), weight = sum(SHIPMT_WGHT), avg_distance = mean(SHIPMT_DIST_ROUTED), prop_h = sum(HAZMAT != "N")/sum(HAZMAT == "N"))

Dest_summary$value[is.na(Dest_summary$value)] <- 0 #replaced NA with 0
Dest_summary$weight[is.na(Dest_summary$weight)] <- 0 # Return to this: why some destinations have a dollar value but no weight information? (services trade? no indication of suppressed values)
```

Example 2: Textiles, Leather, and Articles of Textiles or Leather
```{r}
textiles <- cfs_by_dest[cfs_by_dest$SCTG == 30 ,]

Dest_summary_t <- summarize(textiles, t_value = sum(SHIPMT_VALUE), t_weight = sum(SHIPMT_WGHT), t_avg_distance = mean(SHIPMT_DIST_ROUTED), t_prop_h = sum(HAZMAT != "N")/sum(HAZMAT == "N"))

Dest_summary_t$t_value[is.na(Dest_summary_t$t_value)] <- 0 
Dest_summary_t$t_weight[is.na(Dest_summary_t$t_weight)] <- 0
```

Note: proportion of hazardous material can be weighted by value or weight of shipments


Combine
```{r}
Summary <- merge(Dest_summary_t, Dest_summary, by = "DEST_CFS_AREA")
```


Write CVS
```{r eval=FALSE}
write.csv(Summary, "cfs_r.csv")
```

# Objectives

Interactive map (use Shiny):

* Pounds per CFS area (orgin & destination)

* Dollar value per CFS area (orgin & destination)

* Network Centrality and Power per CFS area

* Show all of above by industry, trasportation mode, type of good transported, and whether material is hazardous.

* Show percentages of values suppressed

Further steps: characteristics of counties can be associated to centrality of CFS areas, by industry. Possible values: economic indicators of state, migration / population growth, employment, etc.

Further steps: combine layers to illustrate mutiple associations (shipments of hazardous material and points with toxic release emission points - from EPA), comparison of received vs sent, value & distance travelled. Note: Comparison maps can be displayed side by side or with different symbols on the same panel.
